---
title: "Replot Drug Curves"
author: "Sara Gosline, Jessica Bade"
date: "10/19/2021"
output: html_document
---

For each of the drugs and models of interest we would like to plot all of the dose response curves in the same cformat so that they can be combined.

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(remotes)
if(!require(mpnstXenoModeling)){
  remotes::install_github('sgosline/mpnstXenoModeling')
  library(mpnstXenoModeling)
}

if(!require('DT')){
  install.packages("DT")
  library(leapr)
}

library(dplyr)
library(tidyr)
library(ggrepel)
```

## Load drug data and expression data

```{r mutation data,message=FALSE,warning=FALSE,echo=FALSE}

if(!exists("dataLoaded")){
  dataLoaded=TRUE
  mpnstXenoModeling::loadPDXData()
  mtDrugData <- loadMicrotissueDrugData()

}
```

## Plot drugs

We now want to plot drug performance across samples of interest, so we can do a side by side of drugs and their combinations.

```{r drug list}

##these are the drugs we want to evaluate!!!
drug.list<-list(MirdaTrab=c('Mirdametinib','Mirdametinib;Trabectedin','Trabectedin'),
                OlaTrab=c('Olaparib','Olaparib;Trabectedin','Trabectedin'),
                ShpRibo=c('SHP099','Ribociclib;SHP099','Ribociclib'),
                ShpTram=c('SHP099','Trametinib','SHP099;Trametinib'))
drug.list.inc<-lapply(drug.list,tolower)
eval.drugs.pdx<-c('mirdametinib','mirdametinib+trabectedin','olaparib','trabectedin','ribociclib','trametinib','trametinib+ribociclib')

eval.samps<-c('MN2','WU-225','JHU 2-079c','JHU 2-031','MN1','WU-368')
eval.samps.inc <- c('MN-2', 'JH-2-079c') #other samples available: JHU 2-002, JHU 2-103
```

### Microtissue Metrics

The first thing we need to do is establish good metrics for evaluating drug sensitivity. We experiment with a few here.

We are focusing on just 3 drugs and 2 combinations at this stage, across 4 PDX/MT/cell line combinations.

### Microtissue Dose Response

```{r plot MT dose response, warning=FALSE, message=FALSE}
library(cowplot)
mtDrugData<-loadMicrotissueDrugData()
##commented this out as it doesn't work. 
#lapply(names(drug.list), function(drugGroup) {
#  mt.subs <- subset(mtDrugData, CellLine %in% eval.samps) %>%
#    subset(Drug %in% drug.list[[drugGroup]])
  # 
  # samp.plots <- lapply(eval.samps, function(samp)
  #   plotMTTreatmentByDrug(mt.subs, samp))
  # 
  # ggsave(
  #   paste0('SampleMT', drugGroup, 'plots.png'),
  #   cowplot::plot_grid(plotlist = samp.plots, ncol = 2),
  #   height = 12,
  #   width = 12
  # )


#  drug.plots<-lapply(intersect(drug.list[[drugGroup]],mtDrugData$DrugCol),
  #                   function(drug) plotMTTreatmentBySample(mt.subs,drug))

 # ggsave(paste0('DrugMT',drugGroup,'plots.png'),
  #       cowplot::plot_grid(plotlist = drug.plots,ncol=2),height=12,width=12)
#})



```

### PDX Dose Response

```{r plot PDX dose response, warning=FALSE, message=FALSE}

drugData<-loadPDXDrugData()

plots <- drugData %>%
  subset(Sample %in% eval.samps) %>%
  subset(drug %in% eval.drugs.pdx) %>%
  group_by(Sample, batch, .drop = F) %>%
  group_map( ~ plotPDXTreatmentBySample(.x), .keep = TRUE)

ggsave('SamplePDXplots.png',
       cowplot::plot_grid(plotlist = plots, ncol = 2))

```

### Incucyte Dose Response


```{r}

icyteData<-loadIncucyteData()
lapply(names(drug.list.inc),function(drugGroup){
  inc.subs<-subset(icyteData,model_system_name%in%eval.samps.inc)%>%
  subset(experimentalCondition%in%drug.list.inc[[drugGroup]])
  inc.subs
  samp.plots<-lapply(eval.samps.inc,function(samp) 
    plotIncTreatmentByDrug(inc.subs,samp))

  ggsave(paste0('SampleInc',drugGroup,'plots.png'),
         cowplot::plot_grid(plotlist = samp.plots,ncol=2),height=6,width=12)

#  drug.plots<-lapply(intersect(drug.list[[drugGroup]],mtDrugData$DrugCol),
  #                   function(drug) plotMTTreatmentBySample(mt.subs,drug))

 # ggsave(paste0('DrugMT',drugGroup,'plots.png'),
  #       cowplot::plot_grid(plotlist = drug.plots,ncol=2),height=12,width=12)
})


```

### Cell Line Dose Response
