---
title: "01-molecular-drug-test"
author: "Sara Gosline"
date: "7/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mpnstXenoModeling)
library(dplyr)
```

Here we have an example of using the `Xeva` package to search for any molecular indicates of drug response in the MPNST PDX data.

## Load PDX Data

PDX data is available on [this Synapse project page](https://www.synapse.org/#!Synapse:syn21984813/wiki/602362) that collates available PDX and tumor model data across the project.

For this project we use the `Xeva` xenograft visualization and analysis tool to test out initial hypotheses of gene expression and mutation correlation.

```{r Build PDX Data Models, echo=FALSE,message=FALSE,warning=FALSE}

loadPDXData()

```
We currently have rnaSeq and drug-sensitivity for 6 PDX models. We can plot the tumor volume data is shown here.

```{r tumor data, echo=FALSE,message=FALSE,warning=FALSE}

#batches=batchInfo(all.pdxs)
library(ggplot2)
plotBySample<-function(dt){
  #print(dt$Sample)
  sample=dt$Sample[1]
  batch=dt$batch[1]
  #print(sample)
  tt<-dt%>%dplyr::select(time,volume,drug)%>%distinct()
  tm <-tt%>%group_by(time,drug)%>%summarize(mvol=median(volume),minVol=median(volume)-sd(volume),maxVol=median(volume)+sd(volume))%>%distinct()
  
  ggplot(tm,aes(x=time,y=mvol,ymin=minVol,ymax=maxVol,col=drug,fill=drug))+geom_line()+ggtitle(sample)+geom_ribbon(alpha=0.25)
}

plots <-drugData%>%
  group_by(Sample,batch,.drop=F)%>%
  group_map(~plotBySample(.x),.keep=TRUE)
 # lapply(batches,function(x) plotGrowthForBatch(all.pdxs,x))
#mods <-sapply(batches,function(x) unlist(strsplit(x,split='_'))[1])

pdf('allTumorVolume.pdf',height=8,width=12)
#res<-lapply(unique(mods),function(x)
#  cowplot::plot_grid(plotlist=plots[grep(x,mods)]))
res=lapply(plots, print)
dev.off()
#res<-lapply(unique(mods),function(x)
#  cowplot::plot_grid(plotlist=plots[grep(x,mods)]))
res=lapply(plots, print)


```

## Visualize mutational data and gene expression


Now we can look at the mutational data in more detail
```{r mut data,echo=FALSE}
 mutMat<-varData%>%mutate(AD=as.numeric(AD))%>%
  dplyr::select(specimenID,Symbol,AD)%>%#,tranche))%>%
  subset(Symbol!="")%>%
  tidyr::pivot_wider(names_from=specimenID,values_from=AD,values_fn=list(AD=mean),
                     values_fill=0.0)%>%
    tibble::column_to_rownames('Symbol')

annotes <- varData%>%
  dplyr::select(Sample,specimenID,Age,Sex,MicroTissueQuality,Location,Size)%>%
  distinct()%>%
  tibble::column_to_rownames('specimenID')%>%
  mutate(Age=as.numeric(Age),Size=as.numeric(Size),MicroTissueQuality=unlist(MicroTissueQuality))
library(pheatmap)
  #ssinds=grep('_',varData$specimenID)
  #varData$specimenID[-ssinds]<-paste(varData$specimenID[-ssinds],'_',sep='')
  #annotes<-varData%>%
  #    dplyr::select(specimenID)%>%#,tranche)%>%
  #  tidyr::separate(specimenID,into=c('patient','sample'),sep='_',remove=F,fill='left')%>%
  #  distinct()%>%
  #  tibble::column_to_rownames('specimenID')%>%
  #  filter(sample!='normal')#

  mutMat<-mutMat[,intersect(rownames(annotes),colnames(mutMat))]
 pheatmap(log10(0.01+mutMat),clustering_distance_cols = 'correlation',cellwidth = 10,annotation_col = annotes,labels_row = rep("",nrow(mutMat)),labels_col=rep("",ncol(mutMat)))
 
 pheatmap(log10(0.01+mutMat),clustering_distance_cols = 'correlation',cellwidth = 10,annotation_col = annotes,labels_row = rep("",nrow(mutMat)),labels_col=rep("",ncol(mutMat)),filename='allMutations.pdf')
```



We then select for those mutations that exist in more than one sample. And then add back TP53.
```{r main mutations,echo=FALSE}
topMuts=filter(varData,AD>0)%>%
  subset(!is.na(Symbol))%>%
  subset(Symbol!="")%>%
    group_by(Symbol)%>%
    summarize(nSamps=n_distinct(individualID))%>%
    filter(nSamps>2)%>%
    dplyr::select(Symbol)

topMuts<-union("TP53",topMuts$Symbol)

 pheatmap(log10(0.01+mutMat[topMuts,rownames(annotes)]),cellwidth = 10,cellheight=10,annotation_col = annotes,clustering_method='ward.D2')
 
 pheatmap(log10(0.01+mutMat[topMuts,]),cellwidth = 10,cellheight=10,annotation_col = annotes,filename='recMutations.pdf',clustering_method='ward.D2')

```

We can also plot the gene expression across highly variable genes.
```{r mutated gene expression,echo=FALSE,warning=FALSE}
 rnamat<-rnaSeq%>%
  subset(!is.na(Symbol))%>%
  subset(!is.na(totalCounts))%>%
  dplyr::select(-c(individualID,species,sex,zScore,experimentalCondition))%>%
  tidyr::pivot_wider(names_from=specimenID,values_from=totalCounts,
                     values_fn=list(totalCounts=mean),values_fill=0.0)%>%
    tibble::column_to_rownames('Symbol')

rownames(rnaSeq)<-c()
 annotes<-rnaSeq%>%
      dplyr::select(specimenID,individualID,sex)%>%
   distinct()%>%
    tibble::column_to_rownames('specimenID')

 #topMuts<-setdiff(topMuts,'TTN')
pheatmap(log10(0.01+rnamat[intersect(topMuts,rownames(rnamat)),]),
          clustering_distance_cols = 'correlation',
          cellwidth = 10,
       #  annotation_col = annotes,
         cellheight=10,
         labels_col=rep("",ncol(mutMat)))

 pheatmap(log10(0.01+rnamat[intersect(topMuts,rownames(rnamat)),]),
          clustering_distance_cols = 'correlation',
          cellwidth = 10,
        # annotation_col = annotes,
         cellheight=10,
         labels_col=rep("",ncol(mutMat)),filename='countsOfMutatedGenes.pdf')
 
 
```

