---
title: "04-mutation-diversity"
author: "Sara Gosline"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(remotes)
if(!require(mpnstXenoModeling)){
  remotes::install_github('sgosline/mpnstXenoModeling')
  library(mpnstXenoModeling)
}
library(dplyr)
```

Here we have an example of using the `Xeva` package to search for any molecular indicates of drug response in the MPNST PDX data.

## Load PDX Data

PDX data is available on [this Synapse project page](https://www.synapse.org/#!Synapse:syn21984813/wiki/602362) that collates available PDX and tumor model data across the project.

```{r Build PDX Data Models, echo=FALSE,message=FALSE,warning=FALSE}

loadPDXData()


```



## Mutational data


Now we can look at the mutational data in more detail
```{r mut data,echo=FALSE}
 mutMat<-varData%>%
  mutate(AD=as.numeric(AD))%>%
  subset(Symbol!="")%>%
  dplyr::select(-c(individualID,specimenID,Sample,Age,Sex,MicroTissueQuality,Location,Size,Clinical.Status))%>%
  distinct()%>%
  tidyr::pivot_wider(names_from=synid,values_from=AD,values_fn=list(AD=mean),
                     values_fill=list(AD=0.0))%>%
    tibble::column_to_rownames('Symbol')

library(pheatmap)
  annotes<-varData%>%
    dplyr::select(c(synid,Sample,Age,Sex,MicroTissueQuality,Location,Size,Clinical.Status))%>%
    mutate(MicroTissueQuality=unlist(MicroTissueQuality))%>%
    mutate(`Clinical Status`=unlist(Clinical.Status))%>%
    dplyr::select(-Clinical.Status)%>%
    #tidyr::separate(specimenID,into=c('patient','sample'),remove=TRUE)%>%
    distinct()%>%
    tibble::column_to_rownames('synid')#%>%
#    dplyr::filter(sample!='normal')
  
  mutMat<-mutMat[,rownames(annotes)]
 pheatmap(log10(0.01+mutMat),clustering_distance_cols = 'correlation',cellwidth = 10,annotation_col = annotes,labels_row = rep("",nrow(mutMat)),labels_col=rep("",ncol(mutMat)))
 
 pheatmap(log10(0.01+mutMat),clustering_distance_cols = 'correlation',cellwidth = 10,annotation_col = annotes,labels_row = rep("",nrow(mutMat)),labels_col=rep("",ncol(mutMat)),filename='allMutations.pdf')
```

## Top mutations


We then select for those mutations that exist in more than one sample. And then add back TP53.
```{r main mutations,echo=FALSE}
topMuts=dplyr::filter(varData,AD>0)%>%
    group_by(Symbol)%>%
    summarize(nSamps=n_distinct(individualID))%>%
    dplyr::filter(nSamps>1)%>%
    dplyr::select(Symbol)
topMuts<-union("TP53",topMuts$Symbol)

 pheatmap(log10(0.01+mutMat[topMuts,]),cellwidth = 10,cellheight=10,annotation_col = annotes,labels_col=rep("",ncol(mutMat)),clustering_method='ward.D2', clustering_distance_cols = 'correlation')
 pheatmap(log10(0.01+mutMat[topMuts,]),cellwidth = 10,cellheight=10,annotation_col = annotes,filename='recMutations.pdf',clustering_method='ward.D2', clustering_distance_cols = 'correlation')

```

## Selected mutations
Below is a list of custom genes we want to evaluate.

```{r select gene mutations,echo=FALSE,warning=FALSE}
genelist=c('NF1','CDKN2A','TP53','EED','SUZ12','EGFR','PDGFRA','MET','BRAF',
           'TYK2','PIK3CA','AURKA2','ATRX','TSC1','TSC2','NF2')
missed.genes<-setdiff(genelist,rownames(mutMat))
print(paste('Missing mutations in genes',paste(missed.genes,collapse=',')))

missed.mat<-matrix(0,nrow=length(missed.genes),ncol=ncol(mutMat),
                   dimnames=list(missed.genes,colnames(mutMat)))

new.mat<-rbind(mutMat[intersect(genelist,rownames(mutMat)),],missed.mat)

 pheatmap(log10(0.01+new.mat),cellwidth = 10,cellheight=10,annotation_col = annotes,labels_col=rep("",ncol(mutMat)),clustering_method='ward.D2')
 pheatmap(log10(0.01+new.mat),cellwidth = 10,cellheight=10,annotation_col = annotes,filename='selectedMutations.pdf',clustering_method='ward.D2')
```

