---
title: "Fig1_SomaticVariantExpression"
author: "Sara Gosline"
date: "9/13/2021"
output:  
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(remotes)

if(!require(mpnstXenoModeling)){
  remotes::install_github('sgosline/mpnstXenoModeling')
  library(mpnstXenoModeling)
}

if(!require('leapr')){
  remotes::install_github('biodataganache/leapR')
  library(leapr)
}
library(dplyr)
```


## Mutation data

Load the data from the R package.

```{r mutation data,echo=FALSE,warning=FALSE,include=FALSE,message=FALSE}
mpnstXenoModeling::loadPDXData()

```


## Mutation heatmaps

We now can select specific mutations of interest and put them in a heatmap

```{r mut data, results='asis',warning=FALSE}
 mutMat<-varData%>%
  mutate(AD=as.numeric(AD))%>%
  subset(Symbol!="")%>%
  dplyr::select(-c(individualID,specimenID,Sample,Age,Sex,MicroTissueQuality,Location,Size,Clinical.Status))%>%
  distinct()%>%
  tidyr::pivot_wider(names_from=synid,values_from=AD,values_fn=list(AD=mean),
                     values_fill=list(AD=0.0))%>%
    tibble::column_to_rownames('Symbol')

library(pheatmap)
  annotes<-varData%>%
    dplyr::select(c(synid,Sample,Age,Sex,MicroTissueQuality,Location,Size,Clinical.Status))%>%
    mutate(MicroTissueQuality=unlist(MicroTissueQuality))%>%
    mutate(`Clinical Status`=unlist(Clinical.Status))%>%
    dplyr::select(-Clinical.Status)%>%
    #tidyr::separate(specimenID,into=c('patient','sample'),remove=TRUE)%>%
    distinct()%>%
    tibble::column_to_rownames('synid')#%>%
#    dplyr::filter(sample!='normal')
  
  mutMat<-mutMat[,rownames(annotes)]
  
   
 pheatmap(log10(0.01+mutMat),clustering_distance_cols = 'correlation',cellwidth = 10,annotation_col = annotes,labels_row = rep("",nrow(mutMat)),labels_col=rep("",ncol(mutMat)),filename='allMutations.pdf')

pheatmap(log10(0.01+mutMat),clustering_distance_cols = 'correlation',cellwidth = 10,annotation_col = annotes,labels_row = rep("",nrow(mutMat)),labels_col=rep("",ncol(mutMat)))
```

Here we see that some samples are more mutated than others. But really, we are only interested in a handful of mutations. 

```{r selected muts, warning=FALSE,results='asis'}
genelist=c('NF1','CDKN2A','TP53','EED','SUZ12','EGFR','PDGFRA','MET','BRAF',
           'TYK2','PIK3CA','AURKA2','ATRX','TSC1','TSC2','NF2')
missed.genes<-setdiff(genelist,rownames(mutMat))
print(paste('Missing mutations in genes',paste(missed.genes,collapse=',')))

 
  smutMat<-mutMat[intersect(genelist,rownames(mutMat)),rownames(annotes)]
  
 pheatmap(log10(0.01+smutMat),cellheight=10,cellwidth = 10,annotation_col = annotes,labels_col=rep("",ncol(smutMat)),filename='selectedMutations.pdf')
 
 pheatmap(log10(0.01+smutMat),cellheight=10,cellwidth = 10,annotation_col = annotes,labels_col=rep("",ncol(smutMat)))
```


Lastly we want to add any recurrent mutations that could be of interest to our analysis.

```{r recurrent mutations,results='asis',fig.keep='first'}
topMuts=dplyr::filter(varData,AD>0)%>%
  subset(!is.na(Symbol))%>%
    group_by(Symbol)%>%
    summarize(nSamps=n_distinct(individualID))%>%
    dplyr::filter(nSamps>2)%>%
    dplyr::select(Symbol)

 
  rmutMat<-mutMat[intersect(union(topMuts$Symbol,genelist),rownames(mutMat)),rownames(annotes)]
  
 pheatmap(log10(0.01+rmutMat),cellheight=10,cellwidth = 10,annotation_col = annotes,labels_col=rep("",ncol(rmutMat)),filename='selectedAndRecMutations.pdf')
 
 pheatmap(log10(0.01+rmutMat),cellheight=10,cellwidth = 10,annotation_col = annotes,labels_col=rep("",ncol(rmutMat)))

```

## RNA signatures of mutations

For each mutation in the `rmutMat` matrix shown above, we  can separate out patients that have the mutation and do not. Then we can compare differential expression using the tools described previously (the third markdown):

1. Group the samples into two groups, then compute the log2fold change and significance via the linear model
2. Plot the significant gene in a heatmap
3. Evaluate the functional enrichment of the significant genes only
4. Evaluate the functional enrichment of the _ranking_ of all genes using GSEA. This is particularly useful when we have no significantly different genes.

```{r diffex by gene,warning=FALSE}

```

