---
title: "Replot Drug Curves"
author: "Sara Gosline"
date: "10/19/2021"
output: html_document
---

For each of the drugs and models of interest we would like to plot all of the dose response curves in the same cformat so that they can be combined.

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(remotes)
if(!require(mpnstXenoModeling)){
  remotes::install_github('sgosline/mpnstXenoModeling')
  library(mpnstXenoModeling)
}

if(!require('leapr')){
  remotes::install_github('biodataganache/leapR')
  library(leapr)
}

if(!require('DT')){
  install.packages("DT")
  library(leapr)
}

library(dplyr)
library(ggrepel)
```


## Load drug data and expression data
```{r mutation data,message=FALSE,warning=FALSE}
mpnstXenoModeling::loadPDXData()


```

## Plot drugs

### Microtissue Metrics

The first thing we need to do is establish good metrics for evaluating drug sensitivity.  We experiment with a few here. 


```{r MT drug plots, message=FALSE, warning=FALSE}

##these are the drugs we want to evaluate!!!
eval.drugs<-c('Mirdametinib','Mirdametinib;Trabectedin','Trabectedin','Olaparib','Olaparib;Trabectedin','mirdametinib','mirdametinib+trabectedin','olaparib','trabectedin')

eval.samps<-c('MN2','WU-225','JHU 2-079','JHU 2-031')



```

We are focusing on just 3 drugs and 2 combinations at this stage, across 4 PDX/MT/cell line combinations.


### Microtissue Dose Response
```{r plot MT dose response, warning=FALSE, message=FALSE}

```


### PDX Dose Response
```{r plot PDX dose response, warning=FALSE, message=FALSE}

#plotPDXTreatmentBySample

```

### Incucyte Dose Response

### Cell Line Dose Response